---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aae.pop.inference

<!-- badges: start -->
![R-CMD-check](https://github.com/aae-stats/aae.pop.inference/actions/workflows/check-standard.yaml/badge.svg)
[![Codecov test coverage](https://codecov.io/github/aae-stats/aae.pop.inference/main/graph/badge.svg)](https://app.codecov.io/github/aae-stats/aae.pop.inference/)
<!-- badges: end -->

`aae.pop.inference` supports inference on population dynamics models implemented
in `aae.pop`. These models are relatively complex, making standard inference
methods (e.g. maximum likelihood, MCMC) challenging, especially as the models
become realistically complex (many population classes, temporal variability). 
`aae.pop.inference` uses a sequential approximate Bayesian computation approach
to support fast inference on these complex models.


## Installation

You can install the released version of aae.pop.inference from 
[GitHub](https://github.com/aae-stats/aae.pop.inference) with:

``` r
remotes::install_github("aae-stats/aae.pop.inference")
```

## Example: estimating carrying capacity and density dependence

This example uses data simulated from a model as a target to estimate
the parameters of this model. This relatively simple example (estimating
parameters of a known model) illustrates how to perform inference on these
models but also highlights a case in which the target parameters are
highly correlated.

```{r basic-example}
# some packages needed
library(aae.pop)
library(aae.pop.inference)

# define a basic population model with aae.pop
nclass <- 5
popmat <- matrix(0, ncol = nclass, nrow = nclass)
popmat[transition(popmat)] <- c(0.3, 0.5, 0.7, 0.8)
popmat[reproduction(popmat)] <- c(0, 1, 4, 10)
dd <- density_dependence(
  funs = ricker(100, exclude = 1),
  masks = reproduction(popmat)
)
pop_fn <- dynamics(matrix = popmat, dd)

# simulate some data from this model to use as a target data set
obs <- simulate(
  pop_fn,
  nsim = 1000,
  args = list(density_dependence = list(theta = 0.4))
)

# define wrapper so we can simulate from the model based on
#   the inference parameters alone
popsim_fn <- function(par) {

  tmp <- update(
    pop_fn,
    density_dependence(
      funs = ricker(par[1], exclude = 1),
      masks = reproduction(popmat)
    )
  )

  sim <- simulate(tmp, nsim = 100, args = list(density_dependence = list(theta = par[2])))

  stat_abundance_trend(sim)

}

# calculate the target from the simulated data set
target <- stat_abundance_trend(obs)

# estimate parameters by comparing the simulation model to
#   the simulated data set with uniform priors on 
#   carrying capacity (U[50, 500]) and the theta parameter
#   of the Ricker density dependence (U[0.01, 1])
abc_est <- inference(
  model = popsim_fn,
  prior = list(
    c("unif", 50, 500),
    c("unif", 0.01, 1)
  ),
  target = target,
  nb_simul = 100,
  progress_bar = FALSE
)

```

Plotting the estimated posteriors for the two parameters suggests relatively broad distributions with 
high uncertainty. However, a plot of these two parameters against one another highlights their close 
relationship, with a near perfect linear correlation between the two.

```{r, echo = FALSE}
# plot the estimated parameters
par(mfrow = c(2, 1), mar = c(5.1, 4.5, 2.1, 0.5))
plot(abc_est, which = 1, xlab = "Carrying capacity")
plot(abc_est, which = 2, xlab = "Carrying capacity")

# and show their correlation
plot(
  abc_est$param[, 2] ~ abc_est$param[, 1], 
  xlab = "Carrying capacity", 
  ylab = "Density dependence parameter",
  las = 1, 
  bty = "l", 
  pch = 16, 
  col = "gray50"
)
```

The `define()` function also provides a helper to set up simulations when 
inference is required for vital rates parameters only (the elements of
the population dynamics matrix). This function can be used to flexibly
perform inference on any or all elements in the vital rates matrix with
fixed or variable priors.

```{r define-example}
# define inference model for the transition and reproduction elements
popsim_vital <- define(
  x = pop_fn,
  masks = list(
    survival = aae.pop::transition(pop_fn$matrix),
    reproduction = aae.pop::reproduction(pop_fn$matrix)
  ),
  nsim = 100
)

# and run it (nb_simul too low and p_acc_min too high for a real example)
pars <- inference(
  model = popsim_vital$model,
  prior = popsim_vital$prior,
  target = popsim_vital$stat(obs),
  nb_simul = 100,   # real values should be closer to 1000
  progress_bar = FALSE,
  p_acc_min = 0.3   # default setting is 0.05
)

# plot the parameter estimates (one for each nonzero parameter)
plot(pars)
```


## Issues

Please leave feedback, bug reports or feature requests at the GitHub [issues page](https://github.com/aae-stats/aae.pop.inference/issues)
