---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aae.pop.inference

<!-- badges: start -->
![R-CMD-check](https://github.com/aae-stats/aae.pop.inference/actions/workflows/check-standard.yaml/badge.svg)
[![Codecov test coverage](https://codecov.io/github/aae-stats/aae.pop.inference/main/graph/badge.svg)](https://app.codecov.io/github/aae-stats/aae.pop.inference/)
<!-- badges: end -->

`aae.pop.inference` supports inference on population dynamics models implemented
in `aae.pop`. These models are relatively complex, making standard inference
methods (e.g. maximum likelihood, MCMC) challenging, especially as the models
become realistically complex (many population classes, temporal variability). 
`aae.pop.inference` uses a sequential approximate Bayesian computation approach
to support fast inference on these complex models.

## Installation

You can install the released version of aae.pop.inference from 
[GitHub](https://github.com/aae-stats/aae.pop.inference) with:

``` r
remotes::install_github("aae-stats/aae.pop.inference")
```

## Example: estimating carrying capacity and density dependence

This example uses data simulated from a model as a target to estimate
the parameters of this model. This relatively simple example (estimating
parameters of a known model) illustrates how to perform inference on these
models but also highlights a case in which the target parameters are
highly correlated.

```{r example}
# some packages needed
library(aae.pop)
library(aae.pop.inference)

# define a basic population model with aae.pop
nclass <- 5
popmat <- matrix(0, ncol = nclass, nrow = nclass)
popmat[transition(popmat)] <- c(0.3, 0.5, 0.7, 0.8)
popmat[reproduction(popmat)] <- c(0, 1, 4, 10)
dd <- density_dependence(
  funs = ricker(100, exclude = 1),
  masks = reproduction(popmat)
)
pop_fn <- dynamics(matrix = popmat, dd)

# simulate some data from this model to use as a target data set
obs <- simulate(
  pop_fn,
  nsim = 1000,
  args = list(density_dependence = list(theta = 0.4))
)

# define wrapper so we can simulate from the model based on
#   the inference parameters alone
popsim_fn <- function(par) {

  tmp <- update(
    pop_fn,
    density_dependence(
      funs = ricker(par[1], exclude = 1),
      masks = reproduction(popmat)
    )
  )

  sim <- simulate(tmp, nsim = 100, args = list(density_dependence = list(theta = par[2])))

  stat_abundance_trend(sim)

}

# calculate the target from the simulated data set
target <- stat_abundance_trend(obs)

# estimate parameters by comparing the simulation model to
#   the simulated data set with uniform priors on 
#   carrying capacity (U[50, 500]) and the theta parameter
#   of the Ricker density dependence (U[0.01, 1])
abc_est <- inference(
  model = popsim_fn,
  prior = list(
    c("unif", 50, 500),
    c("unif", 0.01, 1)
  ),
  target = target,
  nb_simul = 100,
  progress_bar = FALSE
)

```

Plotting the estimated posteriors for the two parameters suggests relatively broad distributions with 
high uncertainty. However, a plot of these two parameters against one another highlights their close 
relationship, with a near perfect linear correlation between the two.

```{r, echo = FALSE}
# plot the estimated parameters
par(mfrow = c(2, 1), mar = c(5.1, 4.5, 2.1, 0.5))
hist(abc_est$param[, 1], main = "", xlab = "Carrying capacity")
hist(abc_est$param[, 2], main = "", xlab = "Density dependence parameter")

# and show their correlation
plot(
  abc_est$param[, 2] ~ abc_est$param[, 1], 
  xlab = "Carrying capacity", 
  ylab = "Density dependence parameter",
  las = 1, 
  bty = "l", 
  pch = 16, 
  col = "gray50"
)
```

## Issues

Please leave feedback, bug reports or feature requests at the GitHub [issues page](https://github.com/aae-stats/aae.pop.inference/issues)
